---
title: "Case Study"
author: "Jamie Ash"
date: "05/31/2022"
output: 
  bookdown::html_document2:  
    self-contained: yes
    theme: paper #cerulean #cosmo #journal #readable
    toc: false
    smooth_scroll: true
    toc_depth: 3
    toc_float: true
    number_sections: false
    fig_caption: yes
    code_folding: hide
    # bibliography: ["geothermal.bib", "r.bib"]
    # csl: apa-5th-edition.csl
    # link-citations: yes
---  

<style type="text/css">  
/* Note: CSS uses C-style commenting. */
h1.title{font-size:22px; text-align:center;}
h4.author{font-size:16px; text-align:center;}
h4.date{font-size:16px; text-align:center;}
body{ /* Normal  */ font-size: 13px}
td {  /* Table   */ font-size: 12px}
h1 { /* Header 1 */ font-size: 16px}
h2 { /* Header 2 */ font-size: 14px}
h3 { /* Header 3 */ font-size: 12px}
.math{ font-size: 10pt;}
.hi{ /* hanging indents */ 
    padding-left:22px; 
    text-indent:-22px;
}
.main-container {
  max-width: 1000px;
  margin-left: auto;
  margin-right: auto;
}
blockquote {  
    padding: 10px 20px;
    margin: 0 0 20px;
    font-size: 12px;
    border-left: 5px solid #eee;
}
code.r{ /* code */ 
       font-size: 12px;
}
pre{/*preformatted text*/ 
    font-size: 12px;
}
p.caption {/* figure captions */ 
    font-size: 1.0em;
    font-style: italic; 
} 
.vscroll-plot {
    width: 800px;
    height: 400px;
    overflow-y: scroll;
    overflow-x: hidden;
}
body {
text-align: justify}
</style>

```{r setup, include=FALSE, message = FALSE}
source("../R/functions.R")
source("../R/libraries.R")
rasterOptions(maxmemory = 120e+10, memfrac = 0.9)
knitr::opts_chunk$set(echo = TRUE)
```

Custom functions 
```{r, message = FALSE}
# input is a boolian raster layer of 0/1 
# output is a data frame with only high values included
# Downlsize reduces the size of the data set
# I may like to simultaniously index the chl signal as well
chl_fronts = function(blooms,
                      signal,
                      downsize = 10){
  ras  = raster::flip(blooms, direction = "x")

  idx  = which(values(blooms) == TRUE) # == TRUE
  
  e    = extent(blooms)
  s    = dim(blooms)
  lons = seq(from = e[1], to = e[2], length = s[1]) 
  lats = seq(from = e[3], to = e[4], length = s[2]) 
  grid = pracma::meshgrid(lons, lats)
  # using the blooms index to subset the signal values
  fronts = data.frame(lons = grid$X[idx], lats = grid$Y[idx], value = signal[idx])
  idx = seq(1, nrow(fronts), downsize)
  fronts = fronts[idx, ]
  fronts
}

# use the apply function to apply the calc mean function for every 8 layers

avestack = function(ras, by = 8, fun){
  t = getZ(ras)
  s      = dim(ras)
  idx_8d = rep(1:ceiling(s[3]/8), 8)
  idx_8d = sort(idx_8d)
  rem    = abs(s[3] - length(idx_8d))
  idx_8d = idx_8d[1:(length(idx_8d)-rem)]
  idx_og = 1:s[3]
  t_idx  = unique(idx_8d)
  idx    = data.frame(idx_8d = idx_8d, idx_og = idx_og)
  idx$idx_8d = as.factor(idx$idx_8d)
  idx = split(idx, idx_8d)
  
  # subset and average every eight days
  tempset = function(x, ras) ras[[x$idx_og]]
  days = lapply(idx, FUN = tempset, ras = ras)
  days = lapply(days, FUN = calc, fun = fun, na.rm = TRUE)
  days = brick(days)
  days = setZ(days, z = t[t_idx], name = "time")
  days
  }

# Writing this as a function to take a date range, a chl anomaly signal, 
# and output an area time-series

# Input is 3D chla raster output is time series vector of bloom area
bloom_area = function(b){
  t    = getZ(b)
  rr   = reclassify(b, cbind(0,NA))
  a    = raster::area(rr)
  temp = raster::mask(a, rr)
  
  km2  = cellStats(temp, stat = "sum", na.rm = TRUE)
  data.frame(area = as.numeric(km2), time= t)
}

# input x is a raster of raw chl
# inpur x is a raster of 0/1 bloom non-bloom
bloom_mag = function(x, b){
  t = getZ(x)
  # Area time series for duration 
  rr   = reclassify(b, cbind(0,NA))
  # find area of each cell + mask cells that matter
  temp = raster::mask(x, rr)
  mag  = cellStats(temp, stat ="max", na.rm = TRUE)
  data.frame(mag = as.numeric(mag), time= t)
}  

basic_theme <- function(x,
                        y, 
                        mar = c(1,1,1,1),
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        dt = TRUE,
                        ylab = "",
                        xlab = ""
){
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  par(mar = mar)
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       )
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  title(main = main,
        cex.lab = 2,
        line= 0.75,
        adj = 0)
  title(ylab = ylab, cex.lab = 1.5, line = 2.5)
  title(xlab = xlab, cex.lab = 1.5, line = 2.5)
}


# 
# area_ts = function(x = chla, sdate = NULL, edate = NULL, stat = "sum") {
#   
#   # Find a bloom not bloom matrix
#   t = getZ(x)
#   u = calc(x, fun = median, na.rm = TRUE)
#   o = calc(x, fun = mad, na.rm = TRUE)
#   blooms = x > u + o
#   blooms = setZ(blooms, z = t, name = "time")
#   
#   # subset the bloom by time
#   blooms  <- crop3d(x     = blooms, 
#                     sdate = sdate,
#                     edate = edate)
#   
#   t <- getZ(blooms)
#   
#   # Area time series for duration 
#   rr <- reclassify(blooms, cbind(0,NA))
#   # find area of each cell + mask cells that matter
#   a  <- raster::area(rr)
#   aa <- raster::mask(a, rr)
#   # take the sum of all the cells area for bloom magnitude
#   area <- as.numeric(cellStats(aa, stat = "sum", na.rm = TRUE)) # m2 to km2?
#   
#   # I'm taking the sum of true/flase values. I want to mask the chla by the 
#   # true/false
#   mag  <- crop3d(x     = chla, 
#                  sdate = sdate,
#                  edate = edate)
#   
#   #mag <- raster::mask(mag, rr)
#   # I'm going to try masking this way so simple. 
#   # multiplying by boolian matrix of blooms
#   mag <- mag * blooms
#   val <- cellStats(mag, stat, na.rm = TRUE)
#   
#   data.frame(area = area, val = val, time = t)
#   # CHL timeseries. layer wise mean the seasonally filtered chl
# }

# polymask <- function(ras, 
#                      x = c(-170, -152, -157, -162, -170, -170), 
#                      y = c(  20,   20,   22,  24.5, 28, 20)){
#   Sr1 = Polygon(cbind(x,
#                       y))
#   Spp = SpatialPolygons(list(Polygons(list(Sr1), "s1")))
#   SpP_ras = raster::rasterize(Spp, ras, getCover=TRUE)
#   SpP_ras[SpP_ras == 1] = NA
#   ras <- raster::mask(ras, SpP_ras)
#   ras
# }
# input is a raster output is a boolian raster of bloom/not bloom

bool = function(x){
  u <- calc(x, fun = median, na.rm = TRUE)
  o <- calc(x, fun = mad, na.rm = TRUE)
  boo <- x > (u + o)
  extent(boo) <- extent(x)
  boo <- setZ(boo, z = getZ(x), name = "time")
  boo
}

vectorize  <- function(x) {
  sdate <- getZ(x)
  # x = raster::flip(x, direction = "x")
  # x = raster::flip(x, direction = "y")
  # 
  e = extent(x)
  x = t(x)
  extent(x) = e
  # 
  x     <- rasterToPoints(x)
  x     <- data.frame(x)
  colnames(x) <- c("lon", "lat", as.character(sdate))
  x     <- reshape2::melt(x, id.vars = c("lat", "lon"))
  colnames(x) <- c("lats", "lons", "time", "val")
  x
}

tsect <- function(x, y, z, xreach = 1, yreach = 1, xlen = 120, ylen = 40){
  #make a max min vector of sla and fsle
  coord = c(max(x, na.rm = TRUE), min(x, na.rm = TRUE), 
            max(y, na.rm = TRUE), min(y, na.rm = TRUE))
  
  # use maxmin vector to create fake x, y vectors (downsized) as meshgrid input
  y_vec <- seq(from = coord[4], to = coord[3], length.out = ylen) 
  # may need to invert to be same length 
  x_vec <- seq(from = coord[2], to = coord[1], length.out = xlen)
  xy_grid <- meshgrid(x_vec, y_vec)
  
  #inisilize a matrix of dim fs_grid[1] filled with NA values
  dims <- dim(xy_grid[[1]])
  z_grid <- matrix(data = NA, 
                   nrow = dims[1], 
                   ncol = dims[2], 
                   dimnames = NULL)
  
  for(iy in 1:dims[1]) {
    for(ix in 1:dims[2]){
      # where in the df is the difference greater than the units
      box_x <- which(abs(x - xy_grid[[1]][iy, ix]) <= xreach)
      box_y <- which(abs(y - xy_grid[[2]][iy, ix]) <= yreach)
      # I think the grids are dif sizes and should be subet differently
      #index vector of both cox_sla and box_fsle as one
      #box <- sort(match(box_y, box_x))
      
      # I think this is the correct way to do this
      box <- box_x[box_x %in% box_y]
      
      z_grid[iy, ix] <- mean(z[box], na.rm = TRUE)
    }
  }
  
  list(z_grid, y_vec, x_vec)
}

jamie_theme <- function(x,
                        y, 
                        line = 0.75,
                        adj = 0,
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        #labels = seq(from = min(x)+5, to   = max(x)-5, by   = 15),
                        labels = round(seq(range(x)[1]+2, range(x)[2]-2, length = 5), 1),
                        at =  round(seq(range(x)[1]+2, range(x)[2]-2, length = 5), 1),
                        dt = TRUE,
                        yaxes = TRUE,
                        xaxes = TRUE,
                        ylab = "",
                        xlab = "",
                        cex.main = 2
){
  
  if(dt == TRUE) labels <- format(as.Date(labels), "%b-%d")
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       #asp = asp,
       )
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  if(xaxes){axis(side = 1,
       at = at,
       labels = labels,
       las = 1, 
       lwd = 2, 
       mgp = c(2, 1, 0), 
       cex.axis = 1.25,
       col = colvect("grey22", alpha = 0.9))}
  if(yaxes){axis(side = 2,
       las  = 2, 
       lwd  = 2, 
       mgp  = c(1, 0.75, 0), 
       cex.axis = 1.25,
       col = colvect("grey22", alpha = 0.9))}
  title(main = main,
        cex.lab = cex.main,
        line = line,
        adj = adj)
  title(ylab = ylab, cex.lab = 1.5, line = 2.5)
  title(xlab = xlab, cex.lab = 1.5, line = 2.5)
}


```

# Data processing

Loading the data using ERDDAP
```{r data, message = FALSE, warning = FALSE}
# Bloom start and stop dates
blooms = read.csv("../data/outfiles/full_sum.csv")
# subseting just 2018 for case study
idx = year(blooms$sdate) == 2018
sdate = as.Date(blooms$sdate[idx]) - 14
edate = as.Date(blooms$edate[idx]) + 30
# lons  = c(-158, -130)
# lats  = c(23, 34)
lons = c(-165, -130)
lats = c( 20,   36)

# should try loading chla data from ERDDAP. Or just anomaly this in one function 
# I need to make sure I'm not saving a fuck tone of files
# I'm uploading the large data set for the anomaly calculation
chl = loadchl(url = "https://upwell.pfeg.noaa.gov/erddap",
              id  = "erdMH1chla1day",
              lat = lats,
              lon = lons,
              sdate = as.Date("2010-01-01"),
              edate = as.Date("2020-02-01"),
              large = TRUE,
              by    = 30)

fields =  c("temp",
            "temp_qc",
            "temp_adjusted", 
            "psal", 
            "psal_adjusted", 
            "pres", 
            "pres_adjusted", 
            "time",
            "cycle_number", 
            "latitude", 
            "longitude",
            "platform_number")

float = loadfloat(sdate  = sdate, 
                  edate = edate, 
                  lons  = lons, 
                  lats  = lats,
                  fields = fields)
gc()
```

Correct oreantation
```{r}
```

Remove Region around Hawaii. CZ fornow 
```{r, eval = TRUE, message = FALSE, warning = FALSE, cache = FALSE}
chl = bufcoast(chl, 
               region = "Hawaiian Islands", 
               path = "../data/infiles/USMaritimeLimitsAndBoundariesSHP")
gc()
```

Calculate CHL anomaly
```{r, message = FALSE, warning = FALSE, cache = FALSE}
# I receive an warning when detrending for cells of all NA values (hawaii)
# chl anom. calc. needs at least one full year to work correctly
chla = anomalize(chl, detrend = FALSE)
chlb = bool(chla)
gc()
```
Subset time-span of interest: 2018 summer bloom
```{r, message = FALSE, warning = FALSE, cache = FALSE}
# I could also just load using errdap. This is faster tho
chl_2018  = timesnip(chl,  as.Date(sdate), as.Date(edate))
chla_2018 = timesnip(chla, as.Date(sdate), as.Date(edate))
chlb_2018 = timesnip(chlb, as.Date(sdate), as.Date(edate))
```

Calculate patch area
```{r, message = FALSE}
# integrated = depth integrated
tsa = bloom_area(chlb_2018)
# magnitude might like the raw CHL
tsm = bloom_mag(chl_2018, chlb_2018)
#tsa = area_ts(chla, sdate = sdate, edate = edate, stat = "mean")
bloom_2018 = data.frame(area = tsa$area, mag = tsm$mag, time = tsa$time)

# cleaning work space
rm(tsa, tsm)
```

Process float data
```{r, message = FALSE, warning = FALSE}
# time-series of the mixed layer depth for the 
# Converting all columns to numeric
float$time = as.numeric(as.Date(float$time))

#float$float_serial_no = as.factor(float$float_serial_no)
float$float_serial_no = as.factor(float$platform_number)

float = data.frame(lapply(float, function(x) as.numeric(as.character(x))))

# removing bad float data
float = subset(float, 
                psal > 3 
                & temp > 5 
                & temp < 45 
                & pres > 0
                & pres < 800)

#float$float_serial_no <- as.numeric(float$float_serial_no)
float$float_serial_no <- as.numeric(float$platform_number)
float$time            <- as.numeric(float$time)

# Identifying bad floats and removing them
badf  = subset(float, temp < 10 & pres < 100)
#badid = is.element(float$float_serial_no, badf$float_serial_no)
badid = is.element(float$float_serial_no, badf$platform_number)
badcy = is.element(float$cycle_number, badf$cycle_number)
float = subset(float, !badid & !badcy)
rm(badcy, badid, badf)

# Identifying bad floats and removing them
badf = subset(float, temp > 20 & pres > 300)
#badid = is.element(float$float_serial_no, badf$float_serial_no)
badid = is.element(float$float_serial_no, badf$platform_number)
badcy = is.element(float$cycle_number, badf$cycle_number)
float = subset(float, !badid & !badcy)

rm(badcy, badid, badf)
```

Calculate density and mixed layer depth
```{r, message = FALSE}
# I changed the region so floats suck

# Density Calculation: function from the oce package and calc. density
float$rho = swRho(salinity = float$psal,
                   temperature = float$temp,
                   pressure = float$pres,
                   latitude = float$latitude,
                   eos = getOption("oceEOS", default = "gsw"))
# I should make sure float factors are not lost to NA values

#float = subset(float, !is.na(float_serial_no))
float = subset(float, !is.na(platform_number))

# Mixed layer depth per profile
#id = unique(float[, c("float_serial_no", "cycle_number")])
id = unique(float[, c("platform_number", "cycle_number")])

mixed = data.frame(matrix(data = NA, 
                           nrow = nrow(id), 
                           ncol = ncol(float)+1, 
                           dimnames = NULL))
colnames(mixed) = colnames(float)

for(i in 1:nrow(id)){
  # oneprof = subset(float, 
  #                   float_serial_no == id$float_serial_no[i] & 
  #                     cycle_number == id$cycle_number[i])
  oneprof = subset(float, 
                   float_serial_no == id$platform_number[i] & 
                   cycle_number == id$cycle_number[i])
  if(nrow(oneprof) > 4) mixed[i, ] = mldchu(ctd = oneprof, x = "rho", y = "pres_adjusted", n = 10)
  }
```

Last minute remove bad data
```{r, message = FALSE}
mix = subset(mixed, !pres == 8)
mix = subset(mix, !pres > 100)
chl_float = subset(float, pres < 40) 
```

Loess running average
```{r,message = FALSE}
# I may want to change tsa to include magnitude. Maybe keep seperate
t     = as.numeric(bloom_2018$time)
curve = loess(mix$pres_adjusted ~ mix$time) 
x_mod = seq(min(t), max(t), length.out = length(t))
mix_mod = predict(curve, newdata =  x_mod, type = "response")
bloom_2018$mld = mix_mod
```

CHL concentration and units conversion
```{r, message = FALSE}
# In mg
bloom_2018$chlcon = (bloom_2018$mld * bloom_2018$mag * bloom_2018$area * 1000000)/1000000000
# cleaning
bloom_2018 = subset(bloom_2018, is.finite(bloom_2018$mag))
bloom_2018 = subset(bloom_2018, is.finite(bloom_2018$chlcon))
head(bloom_2018)

bloom_2018$chlcon = bloom_2018$chlcon/1000
bloom_2018$area = bloom_2018$area/100000
bloom_2018$mag = bloom_2018$mag*1000
```

# Data visualization 

## Timeseries {.tabset}

### Surface Area  
```{r,eval = FALSE,   message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# Bloom Area
curve = loess(bloom_2018$area ~ as.numeric(bloom_2018$time)) 
curve = predict(curve, type = "response")
jamie_theme(x = bloom_2018$time, 
            y = bloom_2018$area, 
            main = "2018 patch surface area", 
            xlab = "Date Time",
            ylab = "Area [km^2e05]",
            dt = TRUE)
points(bloom_2018$time, bloom_2018$area, pch = 20) 
points(bloom_2018$time, curve, pch = 20, col = "red") 
```

```{r,  message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE, echo = FALSE, eval = FALSE}
png(file = "../figs/surface_area_2018.png")
curve = loess(bloom_2018$area ~ as.numeric(bloom_2018$time)) 
curve = predict(curve, type = "response")
jamie_theme(x = bloom_2018$time, 
            y = bloom_2018$area, 
            main = "2018 patch surface area", 
            xlab = "Date Time",
            ylab = "Area [km^2e05]")
points(bloom_2018$time, bloom_2018$area, pch = 20) 
points(bloom_2018$time, curve, pch = 20, col = "red")
dev.off()
```

### Concentration  
```{r, eval = FALSE,  message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# CHL Summed Magnitude
curve_values <- loess(bloom_2018$mag ~ as.numeric(bloom_2018$time)) 
curve_values <- predict(curve_values, type = "response")
jamie_theme(x = bloom_2018$time, 
            y = bloom_2018$mag, 
            main = "2018 chlorophyll concentration", 
            xlab = "Date Time",
            ylab = "CHL [ug/m^3]")
points(bloom_2018$time, bloom_2018$mag, pch = 20) 
points(bloom_2018$time, curve_values, pch = 20, col = "red") 

```

```{r, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE, echo = FALSE, eval = FALSE}
png(file = "../figs/concentration_2018.png")
curve_values <- loess(bloom_2018$mag ~ as.numeric(bloom_2018$time)) 
curve_values <- predict(curve_values, type = "response")
jamie_theme(x = bloom_2018$time, 
            y = bloom_2018$mag*1000, 
            main = "2018 chlorophyll concentration", 
            xlab = "Date Time",
            ylab = "CHL [ug/m^3]")
points(bloom_2018$time, bloom_2018$mag, pch = 20) 
points(bloom_2018$time, curve_values, pch = 20, col = "red") 
dev.off()
```

### Biomass  
```{r,  eval = FALSE, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# CHL Concentration
curve_values <- loess(bloom_2018$chlcon ~ as.numeric(bloom_2018$time)) 
curve_values <- predict(curve_values, type = "response")
jamie_theme(x = bloom_2018$time,
            y = bloom_2018$chlcon,
            main = "Intigrated chlorophyll biomass",
            xlab = "Date Time",
            ylab = "Biomass [kt]")
points(bloom_2018$time, bloom_2018$chlcon, pch = 20)
points(bloom_2018$time, curve_values, pch = 20, col = "red") 
```

```{r, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE, echo = FALSE, eval = FALSE}
png(file = "../figs/biomass_2018.png")
curve_values <- loess(bloom_2018$chlcon ~ as.numeric(bloom_2018$time)) 
curve_values <- predict(curve_values, type = "response")
jamie_theme(x = bloom_2018$time,
            y = bloom_2018$chlcon,
            main = "Intigrated chlorophyll biomass",
            xlab = "Date Time",
            ylab = "Biomass [kt]")
points(bloom_2018$time, bloom_2018$chlcon, pch = 20)
points(bloom_2018$time, curve_values, pch = 20, col = "red") 
dev.off()
```

### All three
```{r, fig.width=5, fig.height=8,message = FALSE, fig.show = "hold",  warning = FALSE, echo = FALSE}
#png(file = "../figs/biomass_2018.png")
layout(matrix(c(1,  
                2,
                3), 
              nrow = 3, 
              ncol = 1, 
              byrow = TRUE),
       widths  = c(1),
       heights = c(1, 1, 1))

# Area
par(mar = c(0, 6, 4, 3))
x = as.numeric(bloom_2018$time)
y = bloom_2018$area
bound = gam(area ~ s(as.numeric(time)), data=bloom_2018)
bound = predict.gam(bound, se.fit = TRUE)
ft = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
jamie_theme(x = x, 
            y = y, 
            dt  =FALSE,
            main = "", 
            xlab = "",
            ylab = expression(Surface ~ Area ~ (km^2 ~ e05)),
            line = 0.75,
            adj = 0,
            yaxes = TRUE,
            xaxes = FALSE)
polygon(c(as.numeric(x), rev(as.numeric(x))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(x, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(x, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(x, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 
box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))

# CHL Concentration
par(mar = c(1.5, 6, 1.5, 3))
x = bloom_2018$time
y = bloom_2018$chlcon
bound = gam(chlcon~ s(as.numeric(time)), data=bloom_2018)
bound = predict.gam(bound, se.fit = TRUE)
ft  = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
curve_values = loess(bloom_2018$chlcon ~ as.numeric(bloom_2018$time)) 
curve_values = predict(curve_values, type = "response")
jamie_theme(x = x,
            y = y,
            dt = FALSE,
            line = 0.75,
            adj = 0.0,
            main = "",
            xlab = "",
            yaxes = TRUE,
            ylab = expression(Chlorophyll ~ Biomass ~ (kt)),
            xaxes = FALSE)
polygon(c(as.numeric(x), rev(as.numeric(x))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(x, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(x, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(x, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y, pch = 20, cex = 1.25) 
box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))

# CHL Summed Magnitude
par(mar = c(4, 6, 0, 3))
y = bloom_2018$mag
x = as.Date(as.numeric(bloom_2018$time))
bound = gam(mag ~ s(as.numeric(time)), data=bloom_2018)
bound = predict.gam(bound, se.fit = TRUE)
ft = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
curve_values = loess(bloom_2018$mag ~ as.numeric(bloom_2018$time)) 
curve_values = predict(curve_values, type = "response")
jamie_theme(x = x, 
            y = y, 
            dt = TRUE,
            ylim = c(100, 700),
            line = -1,
            adj = 0.03,
            main = "", 
            xaxes = TRUE,
            yaxes = TRUE,
            xlab = "Date Time",
            ylab = expression(Chlorophyll ~ (ug/m^3))
            )
polygon(c(as.numeric(x), rev(as.numeric(x))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(x, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(x, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(x, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 
box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
#dev.off()
```

### Mixed Layer Depth
```{r, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# Mixed Layer Depth
jamie_theme(x = mix$time, y = mix$pres_adjusted, 
            ylim = c(80, 0),
            main = "2018 Bloom Mixed Layer Depth", 
            xlab = "Date Time",
            ylab = "Pressure [dbar]")
points(mix$time, mix$pres_adjusted, pch = 20)
points(bloom_2018$time, bloom_2018$mld, col = "red", lwd = 3, pch = 20)
```

```{r, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE, echo = FALSE, eval = FALSE}
png(file = "../figs/mixed_depth_2018.png")
# Mixed Layer Depth
jamie_theme(x = mix$time, y = mix$pres_adjusted, 
            ylim = c(80, 0),
            main = "2018 Bloom Mixed Layer Depth", 
            xlab = "Date Time",
            ylab = "Pressure [dbar]")
points(mix$time, mix$pres_adjusted, pch = 20)
points(bloom_2018$time, bloom_2018$mld, col = "red", lwd = 3, pch = 20)
dev.off()
```

## Argo-floats tracks  
```{r,  message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
# Float Tracks
e = extent(-165, -132, 18, 37)
wdmap = getMap(resolution = "high")

#id <- unique(float[, c("longitude", "latitude", "float_serial_no", "time")])
id <- unique(float[, c("longitude", "latitude", "platform_number", "time")])
#id$float_serial_no <- as.factor(id$float_serial_no)
id$float_serial_no <- as.factor(id$platform_number)
#nb.cols <- length(unique(id$float_serial_no))
nb.cols <- length(unique(id$platform_number))
mycolors <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)
#levels(id$float_serial_no) <- mycolors
levels(id$platform_number) <- mycolors

jamie_theme(x = id$longitude, y = id$latitude, 
            dt = FALSE,
            main = "2018 bloom argo-floats tracks", 
            xlab = "Longitude",
            ylab = "Latitude",
            xlim = e[1:2], 
            ylim = e[3:4])
sz <- scale01(id$time)
points(id$longitude, id$latitude,
       #col = as.character(id$float_serial_no),
       col = as.character(id$platform_number),
       cex = sz*1.5,
       pch = 20) 
plot(wdmap, 
     xlim = e[1:2], 
     ylim = e[3:4], 
     asp = 1, 
     bg = "black", 
     border = "black", 
     col = "black", 
     #wrap = c(-180, 180), 
     add = TRUE)
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
```

```{r, echo = FALSE, eval = FALSE, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
png(file = "../figs/float_tracks_2018.png")
e = extent(-165, -132, 18, 37)
wdmap = getMap(resolution = "high")

id <- unique(float[, c("longitude", "latitude", "float_serial_no", "time")])
id$float_serial_no <- as.factor(id$float_serial_no)
nb.cols <- length(unique(id$float_serial_no))
mycolors <- colorRampPalette(brewer.pal(12, "Paired"))(nb.cols)
levels(id$float_serial_no) <- mycolors

jamie_theme(x = id$longitude, y = id$latitude, 
            dt = FALSE,
            main = "2018 bloom argo-floats tracks", 
            xlab = "Longitude",
            ylab = "Latitude",
            xlim = e[1:2], 
            ylim = e[3:4])
sz <- scale01(id$time)
points(id$longitude, id$latitude,
       col = as.character(id$float_serial_no),
       cex = sz*1.5,
       pch = 20) 
plot(wdmap, 
     xlim = e[1:2], 
     ylim = e[3:4], 
     asp = 1, 
     bg = "black", 
     border = "black", 
     col = "black", 
     #wrap = c(-180, 180), 
     add = TRUE)
box(which = "plot", lty = "solid", lwd = 3, col = colvect("grey22", alpha = 0.9))
dev.off()
```

## Transect plots
```{r, eval = FALSE, message = FALSE, fig.show = "hold", out.width = "90%",  warning = FALSE}
col <- colorRampPalette(c("purple", "blue", "cyan", "white", "yellow", 
                          "orangered", "red3"))(50)
#col <- cmocean("haline")(50)
image(xvect, yvect, zmat,
      main = "",
      xlab = "",
      ylab = "",
      #col = col,
      ylim = c(200, 800),
      axes = FALSE)
box(which = "plot", lty = "solid", lwd = 3, col = "grey25")
axis(side = 1, 
     las = 1, 
     lwd = 2,
     mgp = c(2, 1, 0),    
     at = seq(from = min(xvect) + 8, to = max(xvect)-5, by = 15),
     labels = format(as.Date(seq(from = min(xvect)+5, 
                                 to = max(xvect)-5, by = 15)), "%b-%d"),
     cex.axis = 1.25)
axis(side = 2, 
     las = 2, 
     lwd = 2, 
     at = seq(from = 0, to = 800, by = 200),
     labels = sort(seq(from = 0, to = 800, by = 200), decreasing = TRUE),
     mgp = c(1, 0.75, 0), 
     cex.axis = 1.25)
title(main = list("Temperature timeseries of 2018 bloom"), 
      line = 0.75,  
      adj = 0, 
      cex.main = 1.5)
title(xlab = "Date time", line = 2.6, cex.lab = 1.5)
title(ylab = "Pressure [dbar]", line = 2.6, cex.lab = 1.5)
```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

```{r, message = FALSE, fig.show = "hold", out.width = "95%",  warning = FALSE}

```

land of lost code
```{r, eval = FALSE}
# Mesh-grid transect
# tmat = tsect(x = float$time, 
#              y = float$pres, 
#              z = float$temp, 
#              xreach = 2, 
#              yreach = 2,
#              xlen = 120, 
#              ylen = 40)
# 
# yvect = tmat[[2]]
# xvect = tmat[[3]]
# zmat  = apply(tmat[[1]], 2, rev)
# zmat  = flip.matrix(t(zmat))

# Using cloud free reproceed data long term data set
# https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-chl-multi_cci-l3-chl_4km_daily-rep-v02?
# https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_interpolated_4km_daily-rep?
# url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_4km_monthly-rep?"
# chl = loadsat(var = "CHL",
#               sdate = as.Date("2018-01-01"),
#               edate = as.Date("2019-02-01"),
#               lons = lons,
#               lats = lats,
#               org = "1900-01-01",
#               lat_varid = "lat",
#               lon_varid = "lon",
#               by = 30,
#               url = "https://jash:5.Pellegrino@my.cmems-du.eu/thredds/dodsC/dataset-oc-glo-bio-multi-l4-chl_interpolated_4km_daily-rep?")


# x1 = runif(100, 1, 10)
# y1 = runif(100, 1, 10)
# x2 = runif(50, 1, 10)
# y2 = runif(50, 1, 10)
# idx = vector()
# # should loop through shorter dataset
# for(i in 1:length(x2)) {
#   difx = abs(x1 - x1[i])
#   dify = abs(y1 - y2[i])
#   idx[i]  = which.min(difx - dify)
# }

# # function to subset by layer given an index and rasterstack
# s = dim(chla_2018)
# tempset = function(x, ras) ras[[x]]
# idx = list(1:s[3])
# days_2018 = lapply(idx, FUN = tempset, ras = chla_2018)
# days_2018 = lapply(days_2018, fronts)
# 
# # use 8 daya data
# # input is a table with columns chl, lat, lon. Each table reps one day/layer
# sub_blooms = function(x){
#   u = median(x$chl, na.rm = TRUE) 
#   o = mad(x$chl, na.rm = TRUE)
#   idx = x$chl > u + o 
#   x[idx, ]
#   }
# 
# days_2018 = vectorize(chla_2018)
# colnames(days_2018) = c("lons", "lats", "time", "chl")
# days_2018$days = as.factor(days_2018$time)
# days_2018 = split(xyz, days)
# days_2018 = lapply(days_2018, FUN = sub_blooms)
# 
# # or use cellstats
# days_2018 = cellStats(chla_2018, stat = fronts)

```








