---
title: "submeso_time_series"
author: "Jamie Ash"
date: '2022-07-21'
output: html_document
---

```{r setup, include=FALSE, message = FALSE}
source("functions.R")
source("libraries.R")
source("blooms.R")
rasterOptions(maxmemory = 120e+10, memfrac = 0.9)
knitr::opts_chunk$set(echo = TRUE)
```

functions
```{r}
# input is a boolian raster layer of 0/1 
# output is a data frame with only high values included
# Downsize reduces the size of the data set
# I may like to simultaneously index the chl signal as well

# ------------------------------------------------------------------------------
# input is a raster output is a boolian raster of bloom/not bloom
bool = function(x){
  u <- calc(x, fun = median, na.rm = TRUE)
  o <- calc(x, fun = mad, na.rm = TRUE)
  boo <- x > (u + o)
  extent(boo) <- extent(x)
  boo <- setZ(boo, z = getZ(x), name = "time")
  boo
}

# ---------------------------------------------------------------------------
tableit <- function(x) {
  vectorize  <- function(x) {
    sdate <- getZ(x)
    x     <- rasterToPoints(x)
    x     <- data.frame(x)
    colnames(x) <- c("lat", "lon", as.character(sdate))
    x     <- reshape2::melt(x, id.vars = c("lat", "lon"))
    colnames(x) <- c("lats", "lons", "time", "val")
    x
  }
  x     = lapply(x, FUN = vectorize)
  vals  = lapply(x, FUN = subset, select = "val")
  quord = lapply(x, FUN = subset, select = c("lons", "lats", "time"))
  xyz   = data.frame(quord[1], vals)
  
  xyz$time = as.character(xyz$time)
  xyz$time = as.Date(xyz$time)
  xyz
  
  # s = dim(x)
  # grid = meshgrid(lon, lat)
  # lat  = raster(grid$X)
  # lat  = raster(grid$Y)
  # lat  = rep(lat, s[3])
  # lon  = rep(lon, s[3])
  # lat  = values(raster(lat))
  # lon  = values(raster(lat))
}

# ------------------------------------------------------------------------------
jamie_theme <- function(x,
                        y, 
                        mar = c(1,1,1,1),
                        #asp = 1,
                        ylim = range(y),
                        xlim = range(x),
                        main = "",
                        dt = TRUE,
                        ylab = "",
                        xlab = "",
                        yaxes = FALSE,
                        xaxes = FALSE,
                        line = 0.75,
                        adj = 0.1
                        ){
  if(is.numeric(x)) x <- round(x, 1)
  if(is.numeric(y)) y <- round(y, 1)
  par(mar = mar)
  plot(0,
       ylim = ylim,
       xlim = xlim,
       main = "",
       xlab = "",
       ylab = "",
       axes = FALSE,
       )
  if(xaxes){
    axis(side = 1,
         las = 1, 
         lwd = 2, 
         mgp = c(2, 1, 0), 
         cex.axis = 1,
         col = colvect("grey22", alpha = 0.9))
    }
  if(yaxes){
    axis(side = 2,
         las  = 2, 
         lwd  = 2, 
         mgp  = c(1, 0.75, 0), 
         cex.axis = 1,
         col = colvect("grey22", alpha = 0.9))
    }
  grid(nx = NULL, # X-axis divided in two sections
       ny = NULL, # Y-axis divided in three sections
       lty = 2, 
       col = colvect(c("gray69"), alpha = 0.5), lwd = 1)
  box(which = "plot", 
      lty = "solid", 
      lwd = 3, 
      col = colvect("grey22", alpha = 0.9))
  title(main = main,
        cex.lab = 2,
        line= line,
        adj = adj)
  title(ylab = ylab, cex.lab = 1.5, line = 2.5)
  title(xlab = xlab, cex.lab = 1.5, line = 2.5)
}

```

load data
```{r}
#blooms = read.csv("../data/outfiles/full_sum.csv")
# subseting just 2018 for case study
idx = year(blooms$sdate) == 2018
sdate = as.Date(blooms$sdate[idx]) - 22
edate = as.Date(blooms$edate[idx]) + 22
# lons  = c(-158, -130)
# lats  = c(23, 34)
lons = c(-171, -129)
lats = c( 17,   36)

# I keep having issues with the upwelling server
# I need to make sure I'm not saving a fuck tone of files

# # I'm uploading the large data set for the anomaly calculation
# chl = loadsat(url = "https://upwell.pfeg.noaa.gov/erddap",
#               id  = "erdMH1chla1day",
#               lat = lats,
#               lon = lons,
#               sdate = as.Date("2010-01-08"),
#               edate = as.Date("2019-01-01"),
#               large = TRUE,
#               by    = 365)
# 

# I'm uploading the large data set for the anomaly calculation
chl = loadsat(url = "https://coastwatch.noaa.gov//erddap",
              id  = "noaacwNPPN20S3ASCIDINEOFDaily",
              lat = lats,
              lon = lons,
              parameter = c("all"),
              sdate = as.Date("2018-02-10"),
              edate = as.Date("2022-06-01"),
              large = TRUE,
              by = 30)

fsle = load_nc(path = "C:\\Users\\james\\Desktop\\jamieslife\\data\\infiles\\case_study\\",
               patt = "fsle_po_2018.nc")
gc()

sla = load_nc(path = "C:\\Users\\james\\Desktop\\jamieslife\\data\\infiles\\sla\\",
               patt = "SLA_dt_2018.nc",
              vars = c("latitude", "longitude", "sla", "time"))
gc()
```

fix oreantation
```{r}
# for transposing and flippig the raster 
oreant = function(ras, flip = NULL, t1 = FALSE, t2 = FALSE){
  e    = extent(ras)
  time = getZ(ras)
  if(t1) ras = raster::t(ras)
  if(!is.null(flip)) ras = raster::flip(ras,  direction = flip)
  if(t2) ras = raster::t(ras)
  extent(ras) = e
  ras = setZ(ras, z = time, name = "time")
  ras
  }

fsle = oreant(fsle, flip = "y", t1 = TRUE)
gc()
sla  = oreant(sla, flip = "y", t1 = TRUE)
gc()
chl  = oreant(chl, t1 = TRUE)
gc()
```

Crop all to the same extent
```{r crop}
extent(fsle) = extent(fsle) - c(360, 360, 0 ,0)
extent(sla)  = extent(sla) - c(360, 360, 0 ,0)
e = extent(-165, -131, 23, 34)
fsle = raster::crop(fsle, e)
chl  = raster::crop(chl, e)
sla  = raster::crop(sla,  e)
gc()
```

Calculate a CHL anomaly and index bloom values
```{r}
chla = anomalize(ras = chl, detrend = FALSE)
slaa = anomalize(ras = sla, detrend = FALSE)
chlb = bool(x = chla)
gc()
```

Make FSLE and CHL them the same time domain. Make FSLE same temporal resolution as CHL
```{r}
# rasters area list of rasters like c(ras1, ras2, ras3)
# will cut time down to the min end date and min start date so they same length
# uses the timesnip function
domains = function(rasters){
  z       = lapply(rasters, FUN = getZ)
  zmin    = lapply(z, FUN = min, na.rm = TRUE)
  zmax    = lapply(z, FUN = max, na.rm = TRUE)
  sdate   = max(as.Date(unlist(zmin)))
  edate   = min(as.Date(unlist(zmax)))
  rasters = lapply(rasters, FUN = timesnip, sdate = sdate, edate = edate)
  rasters
  }

match = function(x, y){
  ty = getZ(y)
  ty = as.numeric(ty)
  tx = getZ(x)
  tx = as.numeric(tx)
  
  idx  = which(is.element(tx, ty))
  tx   = subset(tx, is.element(tx, ty))
  x = raster::subset(x, idx)
  x = setZ(x, z = as.Date(tx), name = "time")
  x
}

rasters = domains(rasters = c(fsle, sla, slaa, chlb, chla, chl))

fsle = rasters[[1]]
sla  = rasters[[2]]
slaa = rasters[[3]]
chlb = rasters[[4]]
chla = rasters[[5]]
chl  = rasters[[6]]
rm(rasters)
gc() 

fsle = match(x = fsle, y = chlb)
sla  = match(x = sla,  y = chlb)
slaa = match(x = slaa, y = chlb)
gc()
```

Cut to timespan of bloom from table 1
```{r}
# blooms = read.csv("../data/outfiles/full_sum.csv")
# # subseting just 2018 for case study
# idx = year(blooms$sdate) == 2018
# sdate = as.Date(blooms$sdate[idx]) - 22
# edate = as.Date(blooms$edate[idx]) + 22
edate = edate - 17

chl  = timesnip(chl,  sdate, edate)
chla = timesnip(chla, sdate, edate)
chlb = timesnip(chlb, sdate, edate)
fsle = timesnip(fsle, sdate, edate)
sla  = timesnip(sla,  sdate, edate)
slaa = timesnip(slaa,  sdate, edate)
```

Resample to the size of FSLE
```{r resample}
resample = function(ras, to, method = "bilinear"){
  raster::crs(ras) = "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" 
  t = getZ(to)
  e = extent(to)
  extent(ras) = extent(e)
  ras = raster::resample(ras,  to, method = method)
  extent(ras) = e
  ras = setZ(ras, z = as.Date(t), name = "time") 
  ras
  }

chl  = resample(chl,  to = fsle)
gc()
chla = resample(chla, to = fsle)
gc()
chlb = resample(chlb, to = fsle)
gc()
sla  = resample(sla,  to = fsle)
gc()
slaa = resample(slaa, to = fsle)
gc()
fsle = resample(fsle, to = fsle)
gc()
```

Remove coastal influence: SOmething is wrong here
```{r}
chlb = bufcoast(chlb, 
                region = "Hawaiian Islands", 
                path = "../data/USMaritimeLimitsAndBoundariesSHP")
chla = bufcoast(chla, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
chl = bufcoast(chl, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
fsle = bufcoast(fsle, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
sla  = bufcoast(sla, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
slaa = bufcoast(slaa, 
               region = "Hawaiian Islands", 
               path = "../data/USMaritimeLimitsAndBoundariesSHP")
gc()
```

Create table
```{r}
# this is much simpler code but may take alot of memory
xyz = tableit(c(sla, slaa, fsle, chl, chla, chlb))
colnames(xyz) = c("lon", "lat", "time", "sla", "slaa", "fsle", "chl", "chla", "chlb")

xyz$fsle = xyz$fsle * -1
xyz = xyz[!is.na(xyz$sla),]
xyz = xyz[!is.na(xyz$slaa),]
xyz = xyz[!is.na(xyz$fsle),]
gc()
```

Kcluster by day produce table
```{r, message = FALSE}
kcluster = function(x){
  x$fsle = scale01(x$fsle)
  x$sla  = scale01(x$sla)
  
  us = median(x$sla, na.rm = TRUE)
  uf = median(x$fsle, na.rm = TRUE)
  ms = mad(x$sla, na.rm = TRUE)
  mf = mad(x$fsle, na.rm = TRUE)
  
  cents = rbind(c(us,      uf),
                c(us - ms, uf),
                c(us + ms, uf),
                c(us, uf + mf))
  set.seed(225)
  ktbl = kmeans(x[ ,c("sla", "fsle")], 
                 centers   = cents,
                 iter.max  = 50,
                 nstart    = 100, 
                 algorithm = "Lloyd")
  data.frame(ktbl$cluster)[, 1]
}

# changing sla to slaa funciton should be written better
kcluster = function(x){
  x$fsle = scale01(x$fsle)
  x$sla  = scale01(x$slaa)
  
  us = median(x$slaa, na.rm = TRUE)
  uf = median(x$fsle, na.rm = TRUE)
  ms = mad(x$slaa, na.rm = TRUE)
  mf = mad(x$fsle, na.rm = TRUE)
  
  cents = rbind(c(us,      uf),
                c(us - ms, uf),
                c(us + ms, uf),
                c(us, uf + mf))
  set.seed(225)
  ktbl = kmeans(x[ ,c("slaa", "fsle")], 
                 centers   = cents,
                 iter.max  = 50,
                 nstart    = 100, 
                 algorithm = "Lloyd")
  data.frame(ktbl$cluster)[, 1]
}

# d = as.factor(xyz$time)
# k = lapply(split(xyz, d), FUN = kcluster)
# k = unlist(k)
# xyz$k = as.numeric(k)

# apply to the entire dataset
k = kcluster(xyz)
k = as.numeric(k)
#k = unlist(k) # maybe unessisaacry
xyz$k = k

# if I decide to do an entire timeseries then I will need to remove the seasoanl 
# and climitiligical effect form the SLA. I will also need to apply this before
# subsetting the bloom time. DONE as slaa

gc()
```
perform GAM 
```{r}
library(rgdal)
library(mgcViz)
df  = xyz[!is.na(xyz$chl),]
df  = xyz[!is.na(df$chla),]
df$regions = as.numeric(df$k)
# Converting lat lon to distance
xy   = cbind(df$lat, df$lon)
#utms = rgdal::project(xy, "+proj=utm +zone=4 ellps=WGS84") # need to change zones
utms = rgdal::project(xy, "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") # need to change zones
df$northing = utms[, 1]/1000
df$easting  = utms[, 2]/1000

#xyz = xyz[!is.na(xyz$chla),]
df_b = subset(df, chlb == 1)
df_b$doy = as.numeric(format(df_b$time, '%j'))
df_b$regions = as.factor(df_b$regions)
gc()

# I honestly don't know. Why convert it back?
# xyz.sp = xyz
# coordinates(xyz.sp) = c('easting', 'northing')
# gam_2018 = gam(chla ~ s(doy, by = regions) + s(doy) + s(easting, northing), 
#                data = df_b, 
#                family = gaussian(link = "identity"))

gam_2018 = gam(chla ~ s(doy) + s(easting, northing) + s(regions, k = 4, bs = "re"), 
               data = df_b, 
               family = gaussian(link = "identity"))

gam_2018 = gam(chla ~ regions + s(doy) + s(easting, northing), 
               data = df_b, 
               family = gaussian(link = "identity"))

# gam_2018 = gam(chla ~ s(time, by = regions) + s(easting, northing, k = 50), 
#                 data = xyz, 
#                 family = gaussian(link = "identity"))

# gam_2018 = gam(chla ~ regions + s(time) + s(easting, northing, k = 50), 
#                 data = xyz, 
#                 family = gaussian(link = "identity"))

summary(gam_2018)

visreg(gam_2018)
```
# DATA VISUALIZATION

Plotting the k-clusters as an image
```{r, eval = FALSE}
plot(xyz$slaa, xyz$fsle, col = xyz$k, pch = 20, cex = 0.5)
```

Visualizing the gam
```{r, message = FALSE, fig.show = "hold",  fig.align = "center", fig.height = 8, warning = FALSE}
start = as.Date("2018-06-15")
end   = as.Date("2018-08-25")
#b    = plot(gm_time)
#dp   = as.data.frame(cbind(b[[1]]$x, b[[1]]$fit, b[[1]]$se))
#ind  = which(dp$V2 == max(dp$V2))
#peak = as.Date(dp$V1[ind])

b = plot(gam_2018, all = TRUE)

layout.matrix = matrix(c(1, 2,
                          3, 4), 
                        nrow = 2, 
                        byrow = TRUE)

nf = graphics::layout(layout.matrix, 
                       heights = c(1, 1))

mix  = as.data.frame(cbind(b[[2]]$x, b[[2]]$fit, b[[2]]$se))
sub  = as.data.frame(cbind(b[[3]]$x, b[[3]]$fit, b[[3]]$se))
mesn = as.data.frame(cbind(b[[4]]$x, b[[4]]$fit, b[[4]]$se))
mesp = as.data.frame(cbind(b[[5]]$x, b[[5]]$fit, b[[5]]$se))

time = as.data.frame(cbind(b[[1]]$x, b[[1]]$fit, b[[1]]$se))

#geom_point(data= df_abs, aes(x= log(prd_l_1999), y=prd_gr), size = 0.5, alpha = 0.5) 
#mix    = as.data.frame(cbind(b[[1]]$x, b[[1]]$fit, b[[1]]$se))
title = c("Time", "Mixed", "(-) Mesoscale", "(+) Mesoscale", "Sub-mesoscale")
shadedline(x    = mix$V1, 
           ylim = c(min(mix$V2 - mix$V3), max(mix$V2 + mix$V3)),
           y0 = mix$V2,
           y1 = mix$V2 + mix$V3, 
           y2 = mix$V2 - mix$V3,
           title = "",
           line = 0.7,
           adj  = 0,
           main = paste(title[2]),
           ylab = "CHL Effect [mg/m^3]",
           col  = "grey69",
           xlab = "",
           labels = format(as.Date(mix$V1), "%Y-%m-%d"),
           mar = c(3, 5, 4, 1))
# abline(v = peak, col = colvect("grey22", alpha = 0.5), lty = 1,
#        lwd = 2)
abline(v = start, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)
abline(v = end, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)

#sub    = as.data.frame(cbind(b[[2]]$x, b[[2]]$fit, b[[2]]$se))
title = c("Time", "Mixed", "(-) Mesoscale", "(+) Mesoscale", "Sub-mesoscale")
shadedline(x = sub$V1, 
           ylim = c(min(sub$V2 - sub$V3), max(sub$V2 + sub$V3)),
           y0 = sub$V2,
           y1 = sub$V2 + sub$V3, 
           y2 = sub$V2 - sub$V3,
           title = "",
           line = 0.7,
           adj = 0,
           main = paste(title[3]),
           ylab = "",
           col = "grey69",
           xlab = "",
           labels = format(as.Date(sub$V1), "%Y-%m-%d"),
           mar = c(3, 3, 4, 3))
# abline(v = peak, col = colvect("grey22", alpha = 0.5), lty = 1,
#        lwd = 2)
abline(v = start, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)
abline(v = end, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)

#mesn    = as.data.frame(cbind(b[[3]]$x, b[[3]]$fit, b[[3]]$se))
title = c("Time", "Mixed", "(-) Mesoscale", "(+) Mesoscale", "Sub-mesoscale")
shadedline(x = mesn$V1, 
           ylim = c(min(mesn$V2 - mesn$V3), max(mesn$V2 + mesn$V3)),
           y0 = mesn$V2,
           y1 = mesn$V2 + mesn$V3, 
           y2 = mesn$V2 - mesn$V3,
           title = "",
           line = 0.7,
           adj = 0,
           main = paste(title[4]),
           ylab = "CHL Effect [mg/m^3]",
           col = "grey69",
           xlab = "Month of Year",
           labels = format(as.Date(mesn$V1), "%Y-%m-%d"),
           mar = c(5, 5, 2, 1))
# abline(v = peak, col = colvect("grey22", alpha = 0.5), lty = 1,
#        lwd = 2)
abline(v = start, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)
abline(v = end, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)

#mesp    = as.data.frame(cbind(b[[4]]$x, b[[4]]$fit, b[[4]]$se))
title = c("Time", "Mixed", "(-) Mesoscale", "(+) Mesoscale", "Sub-mesoscale")
shadedline(x = mesp$V1,
           ylim = c(min(mesp$V2 - mesp$V3), max(mesp$V2 + mesp$V3)),
           y0 = mesp$V2,
           y1 = mesp$V2 + mesp$V3,
           y2 = mesp$V2 - mesp$V3,
           title = "",
           line = 0.7,
           adj = 0,
           main = paste(title[5]),
           ylab = "",
           col = "grey69",
           xlab = "Month of Year",
           labels = format(as.Date(mesp$V1), "%Y-%m-%d"),
           mar = c(5, 3, 2, 3))
# abline(v = peak, col = colvect("grey22", alpha = 0.5), lty = 1,
#        lwd = 2)
abline(v = start, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)
abline(v = end, col = colvect("grey22", alpha = 0.5), lty = 1,
       lwd = 2)
```

subsisting each region to be plotted
```{r}
reg   = split(xyz, xyz$k)
mix   = reg[[1]]
mesop = reg[[2]]
meson = reg[[3]]
sub   = reg[[4]]

ass = function(x){
  x = x[!is.na(x)]
  sum(x)/length(x)
}

back = lapply(split(xyz$chlb, xyz$time), FUN = ass)
time = as.Date(names(back))
back = as.numeric(unlist(back))

mix = lapply(split(mix$chlb, mix$time), FUN = ass)
mix = as.numeric(unlist(mix))
mix = mix - back

sub = lapply(split(sub$chlb, sub$time), FUN = ass)
sub = as.numeric(unlist(sub))
sub = sub - back

mesop = lapply(split(mesop$chlb, mesop$time), FUN = ass)
mesop = as.numeric(unlist(mesop))
mesop = mesop - back

meson = lapply(split(meson$chlb, meson$time), FUN = ass)
meson = as.numeric(unlist(meson))
meson = meson - back

regions = data.frame(time = time, back = back, mix = mix, sub = sub, mesop = mesop, meson = meson)
regions$time = as.numeric(regions$time)
```

plotting 
```{r, fig.height = 10, fig.width=5}
layout(matrix(c(1, 
                2, 
                3), 
              nrow = 3, 
              ncol = 1,
              byrow = TRUE),
       widths  = c(1),
       heights = c(1, 1, 1.15))

y = regions$sub
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Fronts", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            xaxes = FALSE,
            yaxes = TRUE,
            mar   = c(1, 5, 2, 3))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 

y = regions$mesop
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Anti-Cyclonic", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            xaxes = FALSE,
            yaxes = TRUE,
            mar = c(1, 5, 1, 3))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 

y = regions$meson
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Cyclonic", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            yaxes = TRUE,
            xaxes = FALSE,
            mar = c(5, 5, 1, 3))
lab = seq(from = min(x), to = max(x), length = 5)
axis(side = 1,
     at     = lab,
     labels = as.character(as.Date(lab)), 
     las = 1, 
     lwd = 2, 
     mgp = c(2, 1, 0), 
     cex.axis = 1,
     col = colvect("grey22", alpha = 0.9))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 
rm(x, y)

# y = regions$back
# x = regions$time
# data = regions
# bound = gam(y ~ s(x), data = data)
# bound = predict.gam(bound, se.fit = TRUE)
# ft   = bound$fit
# lwr = bound$fit - bound$se.fit
# upr = bound$fit + bound$se.fit
# idx = as.numeric(names(ft))
# t  = x[idx]
# jamie_theme(x = x, 
#             y = y, 
#             dt = FALSE,
#             main = "Backround", 
#             ylim = range(y, na.rm = TRUE),
#             xlab = "test",
#             ylab = "Surface Area [km^2e05]",
#             line = -2,
#             adj = 0.05,
#             yaxes = FALSE,
#             xaxes = TRUE)
# polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
# lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
# lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
# lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
# points(x, y,  pch = 20, cex = 1.25) 
```

subsisting each region t be plotted
```{r}
reg   = split(xyz, xyz$k)
mix   = reg[[1]]
mesop = reg[[2]]
meson = reg[[3]]
sub   = reg[[4]]

back = lapply(split(xyz$chla, xyz$time), FUN = mean, na.rm=TRUE)
time = as.Date(names(back))
back = as.numeric(unlist(back))

mix = lapply(split(mix$chla, mix$time), FUN = mean, na.rm=TRUE)
mix = as.numeric(unlist(mix))
mix = mix - back

sub = lapply(split(sub$chla, sub$time),  FUN = mean, na.rm=TRUE)
sub = as.numeric(unlist(sub))
sub = sub - back

mesop = lapply(split(mesop$chla, mesop$time), FUN = mean, na.rm=TRUE)
mesop = as.numeric(unlist(mesop))
mesop = mesop - back

meson = lapply(split(meson$chla, meson$time), FUN = mean, na.rm=TRUE)
meson = as.numeric(unlist(meson))
meson = meson - back

regions = data.frame(time = time, back = back, mix = mix, sub = sub, mesop = mesop, meson = meson)
regions$time = as.numeric(regions$time)
```

plotting
```{r, fig.height = 10, fig.width=5}
layout(matrix(c(1, 
                2, 
                3), 
              nrow = 3, 
              ncol = 1,
              byrow = TRUE),
       widths  = c(1),
       heights = c(1, 1, 1.15))

y = regions$sub
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Fronts", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            xaxes = FALSE,
            yaxes = TRUE,
            mar   = c(1, 5, 2, 3))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 

y = regions$mesop
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Anti-Cyclonic", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            xaxes = FALSE,
            yaxes = TRUE,
            mar = c(1, 5, 1, 3))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 

y = regions$meson
x = regions$time
data = regions
bound = gam(y ~ s(x), data = data)
bound = predict.gam(bound, se.fit = TRUE)
ft   = bound$fit
lwr = bound$fit - bound$se.fit
upr = bound$fit + bound$se.fit
idx = as.numeric(names(ft))
t  = x[idx]
jamie_theme(x = x, 
            y = y, 
            dt = FALSE,
            main = "Cyclonic", 
            ylim = range(y, na.rm = TRUE),
            xlab = "",
            ylab = "Percent Association",
            line = -2,
            adj = 0.05,
            yaxes = TRUE,
            xaxes = FALSE,
            mar = c(5, 5, 1, 3))
lab = seq(from = min(x), to = max(x), length = 5)
axis(side = 1,
     at     = lab,
     labels = as.character(as.Date(lab)), 
     las = 1, 
     lwd = 2, 
     mgp = c(2, 1, 0), 
     cex.axis = 1,
     col = colvect("grey22", alpha = 0.9))
polygon(c(as.numeric(t), rev(as.numeric(t))), c(upr, rev(lwr)), lty = 0, col = "grey")
lines(t, ft,  pch = 20, col = "grey22",  cex = 1.25)
lines(t, lwr, pch = 20, col = "grey40", cex = 1.25)
lines(t, upr, pch = 20, col = "grey40", cex = 1.25)
points(x, y,  pch = 20, cex = 1.25) 
rm(x, y)
 
```





















