---
title: "HW5 MTH 321"
author: "Jamie Ash"
date: "2022-09-15"
output: pdf_document
---

**Question 1:** If *a* and *b* are both odd, then *ab* is odd.  
  *Hypothesis:* *a* and *b* are both odd  
  *Conclusion:* *ab* is odd  
  *Truth value:* True  
  *Proof:* 
  Suppose $a$ and $b$ are both odd.  
  Then $a=2k+1$ and $b = 2l+1$ where $k,l \ in Z$.  
  Likewise $ab=(2k+1)(2l+1) = 4kl + 2k+ 2l + 1$.  
  Let $(4kl + 2L +2k) = 2(2kl + L +k) = 2h$ where $h \in Z$.  
  So $ab = 2h + 1$.  
  Therefore $ab$ is odd.  

**Question 2:** If *a* is even or *b* is even then *ab* is even.  
  *Hypothesis:* *a* is even or *b* is even  
  *Conclusion:* *ab* is even  
  *Truth value:* True  
  *Proof:*  
  Suppose $a$ is even or $b$ is even. We approach this proof by addressing the 
  two relevant cases.   
  **Case 1:** Without loss of generality, suppose $a$ is even and $b$ is even.  
  It follows that $a = 2k$ and $b = 2l$ where $k, l \in Z$.  
  Then $ab = 2k \times2l = 2(2kl) = 2h$ where $h \in Z$.  
  Lastly $ab = 2h$ where $h \in Z$.  
  Therefore $ab$ is even.  
  
  **Case 2:** Without loss of generality, suppose $a$ is even and $b$ is odd.  
  It follows that $a = 2k$ and $b = 2l +1$ where $k, l \in Z$.  
  Then $ab = 2k(2l + 1) = 2(2kl + k) = 2h$ where $h \in Z$.  
  Simply $ab = 2h$ where $h \in Z$.  
  Therefore $ab$ is even.  
  
  Therefore in all cases of $a$ is even or $b$ is even, then $ab$ is even.  

**Question 3:** If $x \in Z$ is odd, then $x^2 - 1$ is divided by 4.   
  *Hypothesis:* $x \in Z$ is odd   
  *Conclusion:* $x^2 - 1$ is divided by 4  
  *Truth value:* True  
  *Counter example:*  I was not sure if this statement is true, so I tested it 
  for a few values of $x \in Z$.  

```{r, cache = FALSE}
# creating integers 
z = seq(from = -1000000000, to = 1000000000, by = 1)
# if x is odd then...
x = 2*z+1
# creating x^2 - 1 
c = x^2 - 1
# c %% 4 returns a vector of the remainders of c / 4
# unique(c %% 4) returns a vector of all the unique values in c %% 4
unique(c %% 4)
# The only unique remainder is 0, so all values in c are divisible by 4

# cleaning the environment
rm(z, x, c)
```

All of the remainders of $\frac{x^2 - 1}{4}$ are 0 when 
[$x \in Z: -1000000000 \le x \le  1000000000$], therefore it's likely that 
**Question 3** is a true statement.

**Question 4:** If $a | b$ then $a^2 | b^2$.  
  *Hypothesis:* $a | b$     
  *Conclusion:* $a^2 | b^2$  
  *Truth value:* True  
  Suppose $a | b$   
  Then $a/b = c$ where $c \in Z$.   
  Furthermore $b^2/a^2$ = $a/b$ * $a/b$ = $c * c$ = $c^2$.    
  If $c \in Z$, then $c^2 \in Z$.  
  So $b^2/a^2 = k$ where $k \in Z$.    
  Therefore $a^2 | b^2$.    

**Question 5:** If $a | c$ and $b | c$ then $ab | c$.   
  *Hypothesis:* $a | c$ and $b | c$   
  *Conclusion:* $ab | c$  
  *Truth value:* False  
  *counter:*    
  Ff $c$ is even and $|c| > 2$, and $b = c$ while $a = c/2$, then $ab > c$.   
  Therefore, in general $ab$ does not divide $c$.  

example:
```{r}
c = 4  
b = 4  
a = 2  

# a|b
b/a

# b|c
c/b

# ab|c
c / (a*b)
```
  
**Question 6:** If $d | a$ and $d | b$ then $d | (a + b)$.   
  *Hypothesis:* $b | a$ and $d | b$   
  *Conclusion:* $d | (a + b)$  
  *Truth value:* True 
  *Proof:*
  Suppose $d | a$ and $d | b$.   
  So (a/d) = k where $k \in Z$.   
  Similarly $(a/d) = n$ where $n \in Z$.    
  Using $(a + b)/b = (a/d) + (b/d)$.   
  Therefore $d | (a + b)$.  

**Question 7:** If $a | r$ and $c | q$ then $ac | cq$.  
  *Hypothesis:* $a | r$ and $c | q$   
  *Conclusion:* $ac | cq$  
  *Truth value:* True  
  *Proof:*
  Suppose $a | r$ and $c | q$.  
  Then $r/a = k$ and $c/q = l$ where $k,l \in A$.  
  So $\frac{r}{a} \times \frac{q}{c} = \frac{rq}{ac} = k \times l$ 
    where $k,l \in Z$.  
  Furthermore $k \times l = h$ where $h \in Z$.  
  Lastly $\frac{rq}{ac} = h \in Z$.  
  Therefore $ac | bd$  

**Question 8:** If $n \in Z$ then $3n^2 + n + 114$ is even.  
  *Hypothesis:* $n \in Z$   
  *Conclusion:* $3n^2 + n + 114$ is even  
  *Truth value:* False  
  *Counter:*  
  Suppose $n \in Z$.  
  Case 1: $n$ is odd, then $3n^2 + n + 114$ is odd.  
  Case 2: $n$ is even, then $3n^2 + n + 114$ is even.    
  Therefore $3n^2 + n + 114$ is not generaly even when $n \in Z$.     

**Question 9:** If $x,y \in R$ satisfy $x < y$, then $x^2 < y^2$.  
  *Hypothesis:* $x,y \in R$ satisfy $x < y$  
  *Conclusion:* $x^2 < y^2$  
  *Truth value:* False  
  *Counter:*

When x is positive and y is negative, $x^2 < y^2$ is false.  
Here are a few cases where $x^2$ and $y^2$ are equal...  
```{r}
x  = seq(from = -1, to = -1000, by = -1)
y  = x * -1
x2 = x^2 
y2 = y^2
# all are FALSE
unique(x2 < y2)
unique(x2 == y2)
```

**Question 10:** If $x \in (0, 4)$, then $\frac{4}{x(4-x)} \ge1$.  
  *Hypothesis:* $x \in (0, 4)$  
  *Conclusion:* $\frac{4}{x(4-x)} \ge1$  
  *Truth value:* False?  
  *Counter:*  
  Does this mean x can be either 0 or 4? If x is 0, then the conclusion has a 0
  in the denominator.   
  
 














































